### 2019-09-25 aistats_paper/sections/discussion.tex

%!TEX root = ../aistats_paper.tex
\section{Discussion}

Our results have shown that both of our proposed methods perform comparably to the existing method of fitting extreme deconvolution models, whilst converging faster.
The results also show that using GPU-based computation speeds up fitting considerably, allowing sublinear scaling of training time with mixture component size $K$.

Further improvements to our approaches are possible.
The original paper presents a method of getting out of local maxima by splitting and merging clusters, with the split-merge criteria evaluated on the whole dataset.
It should be possible to replace the criteria with stochastic estimates, which would permit them to be used with both the SGD and minibatch EM methods.
Our approaches also add more free parameters to be selected, including learning rate and batch size.
This adds scope for hyperparameter optimisation to improve the log-likelihood.

In this preliminary study the SGD method provided the best log-likelihood values, was faster to train, and scaled better with component size $K$.
In addition, we found SGD to be more numerically stable than minibatch-EM during training.
Both minibatch methods will allow us to fit larger models going forwards.