### 2019-11-18 poster/poster.tex
Add poster base.

% Gemini theme
% https://github.com/anishathalye/gemini
%
% We try to keep this Overleaf template in sync with the canonical source on
% GitHub, but it's recommended that you obtain the template directly from
% GitHub to ensure that you are using the latest version.

\documentclass[final]{beamer}

% ====================
% Packages
% ====================

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[size=custom,width=84.1,height=118.9,scale=1.175]{beamerposter}
\usetheme{gemini}
\usecolortheme{edinburgh}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage[numbers]{natbib}
\usetikzlibrary{shapes, arrows, positioning, fit, calc, arrows.meta}
\usetikzlibrary{bayesnet, positioning}
\tikzset{>=latex}

\bibliographystyle{abbrv}

% ====================
% Lengths
% ====================

% If you have N columns, choose \sepwidth and \colwidth such that
% (N+1)*\sepwidth + N*\colwidth = \paperwidth
\newlength{\sepwidth}
\newlength{\colwidth}
\setlength{\sepwidth}{0.0333\paperwidth}
\setlength{\colwidth}{0.45\paperwidth}

\newcommand{\separatorcolumn}{\begin{column}{\sepwidth}\end{column}}

\definecolor{phicolor}{HTML}{0074D9}
\definecolor{gammacolor}{HTML}{FF851B}

% Tikz options
\tikzset{%
  var/.style      = {draw, thick, circle, minimum size=0.6cm, inner sep=0},
  block/.style    = {draw, thick, rectangle, rounded corners},
  customline/.style    = {draw opacity=0.4, line width=1}
}
\tikzstyle{arrow} = [draw,line width=0.8pt, black, >=triangle 45]

\definecolor{b}{rgb}{0, 0, 0}
\definecolor{g}{rgb}{0, 0, 0}
\definecolor{r}{rgb}{0, 0, 0}
\definecolor{m}{rgb}{0, 0, 0}

% ====================
% Title
% ====================

\title{Scalable Extreme Deconvolution}

\author{James A. Ritchie, Iain Murray}

\institute[]{School of Informatics, University of Edinburgh}

% ====================
% Body
% ====================

\begin{document}

\addtobeamertemplate{headline}{} 
{
    \begin{tikzpicture}[remember picture,overlay] 
    \node [anchor=north west, inner sep=3cm] at ([xshift=0.cm, yshift=1.cm]current page.north west)     
    {\includegraphics[height=4.cm]{figures/uoe-informatics-white.png}}; 
    \end{tikzpicture} 
}

\addtobeamertemplate{headline}{} 
{
    \begin{tikzpicture}[remember picture,overlay] 
    \node [anchor=north east, inner sep=3cm] at ([xshift=0.cm, yshift=1.cm]current page.north east)     
    {\includegraphics[height=3.cm]{figures/cdt-inverted.eps}}; 
    \end{tikzpicture} 
}

\addtobeamertemplate{headline}{} 
{
    \begin{tikzpicture}[remember picture,overlay] 
    \node [anchor=north east, inner sep=3cm] at ([xshift=0.cm, yshift=-3.cm]current page.north east)     
    {\includegraphics[height=3.cm]{figures/epsrc_logo.pdf}}; 
    \end{tikzpicture} 
}

\begin{frame}[t]
\begin{columns}[t]
\separatorcolumn

\begin{column}{\colwidth}

    \begin{alertblock}{Summary}
    \end{alertblock}

    \begin{block}{Column 1}
    \end{block}

\end{column}

\separatorcolumn

\begin{column}{\colwidth}

    \begin{block}{Column 2}
    \end{block}

    \begin{block}{Acknowledgements}
    \end{block}

    \begin{block}{References}

    \nocite{*}
    \footnotesize{\bibliography{poster}}

    \end{block}

\end{column}

\separatorcolumn
\end{columns}
\end{frame}

\end{document}